---
title: "R Notebook on parsing travel data from GISAID meta data sources"
output: html_notebook
author: Yuanyuan Wang
date: 05/03/2021
---
### A list of relevant packages 
```{r}
packages <- c("countrycode",
              "ggplot2",
              "stringr",
              "formattable",
              "VennDiagram",
              "zoo",
              "devtools",
              "dplyr",
              "tidyr",
              "data.table",
              "webshot",
              "htmltools")
```
### Install packages (only first time)
```{r}
invisible(lapply(packages, install.packages, character.only = TRUE))
install.packages("quanteda", dependencies = TRUE, INSTALL_opts = '--no-lock')
```
### Load packages (every time)
```{r}
###
invisible(lapply(packages, library, character.only = TRUE))
library("quanteda")
```
### Load functions
```{r}
source('get_WHO_countries.R')
source('process_epi_meta.R')
source('process_pkg_meta.R')
source('parse_travel_notes.R')
source('add_voc_voi_label.R')
source('count_seqs_by_country.R')
source('export_travel_seqs.R')
source('export_formattable.R')
```
### Step 1: Provide 4 input files and files names for output files
two meta files from GISAID, the case count file from WHO, and the list of travel words

```{r}
setwd("~/GISAID_metadata_explorer")
Epi_meta_file <- c("input_metadata_from_genomic_epi.tsv")
Download_pkg_meta_file <- c("input_metadata_from_download_packages.tsv")
WHO_case_file <- c("input_who.csv")
List_of_travel_words <- c("input_list_of_travel_words.csv") 
List_of_voc_voi <- c("input_variant_groups.csv")

Output_file_name_GISAID_travel_related_sequences <- c("GISAID_metadata_travel_related.csv")
Output_file_name_country_count_table_figure <- c("Table_of_countries_submitted_sequences_related_to_travel.jpg")
Output_file_name_country_count_table_csv <- c("Table_of_countries_submitted_sequences_related_to_travel.csv")
```
### Step 2: Process input data files, parse travel notes and add VOI/VOC/Spike protein labels
```{r}
setwd("~/GISAID_metadata_explorer/input_files")
who_countries <- get_WHO_countries(WHO_case_file)
gis <- process_epi_meta(who_countries, Epi_meta_file)
gis2 <- process_pkg_meta(who_countries, Download_pkg_meta_file)
gis3 <- parse_travel_notes(gis1, gis2, List_of_travel_words) 
gis4 <- add_voc_voi_label(gis3, List_of_voc_voi)
setwd("~/GISAID_metadata_explorer")
```
### Step 3: Save count table by country as a csv file and a jpg figure, and write the travel related sequences to a file 
```{r}
travel_seqs <- export_travel_seqs(gis4, Output_file_name_GISAID_travel_related_sequences)# this will also write a file

widget_formattable <- count_seqs_by_country(gis3, Output_file_name_country_count_table_csv)
widget_formattable # print the table to screen

# Export table 
export_formattable(widget_formattable,
                   Output_file_name_country_count_table_figure, 
                   width = 800, height = 1250)
```


